---
title: "autocorrelation parameters Timor temp data"
author: "CKim"
date: "05/01/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Testing models for coARMA autocorrelation structures for mixed effects model

After discussion with Simone Blomberg

need to:
1. AIRMA model of temp data changing AR, diff, and MA values - up to 10
   see p 144 of Cowpertwait
   determine best autoregressive orders for ANOVA of temperature data

```{r read in monthly ave temp data, loggers and satellite}
library(tidyverse)
library(here)
log_crw_month <- read_csv(here('data/combined_log_crw_monthttemp.csv') )

str(log_crw_month)

# model error if factor
log_crw_month$year <- as.factor(log_crw_month$year)
#log_crw_month$month <- as.factor(log_crw_month$month)

log_crw_month$season <- as.factor(log_crw_month$season)
log_crw_month$Method <- as.factor(log_crw_month$Method)
```
```{r filter out time series per logger}
log1 <- (log_crw_month[which(log_crw_month$log_id == 1),])
log2 <- log_crw_month[which(log_crw_month$log_id == 2),]
log3_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 3)])
log4_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 4)])
log5_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 5)])
log6_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 6)])
log7_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 7)])

log1_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 1)])
log2_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 2)])
log3_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 3)])
log4_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 4)])
log5_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 5)])
log6_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 6)])
log7_tstemp <- as.ts(log_crw_month$aveTemp[which(log_crw_month$log_id == 7)])
```

no upward trend so diff not necessary?
```{r plot diffs of logger 1}
plot(log1_tstemp)
plot(diff(log1_tstemp))
plot(diff(log1_tstemp, lag = 2))
plot(diff(log1_tstemp, lag = 3))
plot(diff(log1_tstemp, lag = 4))
plot(diff(log1_tstemp, lag = 5))
plot(diff(log1_tstemp, lag = 6))
plot(diff(log1_tstemp, lag = 7))
plot(diff(log1_tstemp, lag = 8))
plot(diff(log1_tstemp, lag = 9))
plot(diff(log1_tstemp, lag = 10))
plot(diff(log1_tstemp, lag = 11))
plot(diff(log1_tstemp, lag = 12))
```

```{r matrix of log ts}
# matrix of ts temps
log_mat <- cbind(log1_tstemp, log2_tstemp, log3_tstemp, log4_tstemp, log5_tstemp, log6_tstemp, log7_tstemp)
str(log_mat)
log_mat <- as.matrix(log_mat)

diff(log_mat, lag = 1)
```


## Test arima model

```{r test arima model, echo=FALSE}
AIC(arima(log1_tstemp, order = c(8,0,0),
          seasonal = list(order = c(8,0,0), 12)))

AIC(arima(log1_tstemp, order = c(0,0,10),
          seasonal = list(order = c(0,0,10), 12)))

```

```{r call get best arima function}
source(here('functions/get_best_arima.R'))
```

```{r log1 run best arima function, warning = F}
log1_best_arima <- get.best.arima(log1_tstemp, maxord = c(6,0,6,6,0,6))
log1_best_arima

log1_best_arimad0 <- get.best.arima.d0(log1_tstemp, maxord = c(6,6,6,6))
log1_best_arimad0
```

```{r log6 run best arima function, include = F}
log1_best_arima <- get.best.arima.d0(log1_tstemp, maxord = c(6,6,6,6))
log2_best_arima <- get.best.arima.d0(log2_tstemp, maxord = c(6,6,6,6))
log3_best_arima <- get.best.arima.d0(log3_tstemp, maxord = c(6,6,6,6))
log4_best_arima <- get.best.arima.d0(log4_tstemp, maxord = c(6,6,6,6))
log5_best_arima <- get.best.arima.d0(log5_tstemp, maxord = c(5,5,5,5))
log6_best_arima <- get.best.arima.d0(log6_tstemp, maxord = c(5,5,5,5))
log7_best_arima <- get.best.arima.d0(log7_tstemp, maxord = c(6,6,6,6))
```

```{r log7 run best arima function}
log1_best_arima[[3]]
log2_best_arima[[3]]
log3_best_arima[[3]]
log4_best_arima[[3]]
log5_best_arima[[3]]
log6_best_arima[[3]]
log7_best_arima[[3]]
```

best p - 2
best q - 0 or 2
diff = 0

not necessary to diff the time series


```{r repeated measures with lme and coARMA autocorrelation, p = 1, q = 0}
library(nlme)
model.lme.ARMA = lme(aveTemp ~ season * Method , 
                random = ~ 1 | log_id,
               correlation = corARMA(form = ~ month_id|log_id, 
                                    p =1,
                                    q =1),
                data = log_crw_month,
                method = "REML")
library(car)
Anova(model.lme.ARMA, type = 'II')
AIC(model.lme.ARMA)
summary(model.lme.ARMA)

plot((model.lme.ARMA))

library(emmeans)
emmeans(model.lme.ARMA, list(pairwise ~ season:Method))
```
```{r update ARMA and compare AIC}

AIC(model.lme.ARMA)
AIC(update(model.lme.ARMA, correlation = corARMA(p = 2, q = 0)))
AIC(update(model.lme.ARMA, correlation = corARMA(p = 3, q = 0)))
AIC(update(model.lme.ARMA, correlation = corARMA(p = 4, q = 0)))
AIC(update(model.lme.ARMA, correlation = corARMA(p = 5, q = 0)))
AIC(update(model.lme.ARMA, correlation = corARMA(p = 6, q = 0)))

AIC(update(model.lme.ARMA, correlation = corARMA(p = 1, q = 1)))
#AIC(update(model.lme.ARMA, correlation = corARMA(p = 2, q = 1)))
AIC(update(model.lme.ARMA, correlation = corARMA(p = 3, q = 1))) 
AIC(update(model.lme.ARMA, correlation = corARMA(p = 4, q = 1)))
AIC(update(model.lme.ARMA, correlation = corARMA(p = 5, q = 1))) ##
#AIC(update(model.lme.ARMA, correlation = corARMA(p = 6, q = 1))) #

AIC(model.lme.ARMA)
AIC(update(model.lme.ARMA, correlation = corARMA(p = 2, q = 2)))
AIC(update(model.lme.ARMA, correlation = corARMA(p = 3, q = 2)))
AIC(update(model.lme.ARMA, correlation = corARMA(p = 4, q = 2))) 
#AIC(update(model.lme.ARMA, correlation = corARMA(p = 5, q = 2))) #
#AIC(update(model.lme.ARMA, correlation = corARMA(p = 6, q = 2)))

AIC(model.lme.ARMA)
#AIC(update(model.lme.ARMA, correlation = corARMA(p = 2, q = 3)))
AIC(update(model.lme.ARMA, correlation = corARMA(p = 3, q = 3)))
#AIC(update(model.lme.ARMA, correlation = corARMA(p = 4, q = 3)))
#AIC(update(model.lme.ARMA, correlation = corARMA(p = 5, q = 3)))
#AIC(update(model.lme.ARMA, correlation = corARMA(p = 6, q = 3)))
```

```{r 'ARMA with p = 3, q = 1'}
ARMA.31 <- lme(aveTemp ~ season * Method , 
                random = ~ 1 | log_id,
               correlation = corARMA(form = ~ month_id|log_id, 
                                    p =3,
                                    q =1),
                data = log_crw_month,
                method = "REML")
   update(model.lme.ARMA, correlation = corARMA(p = 3, q = 1))
AIC(ARMA.31)

summary(ARMA.31)
Anova(ARMA.31)

plot(ARMA.31)

emmeans(ARMA.31, list(pairwise ~ season:Method))
```

```{r autocorrelation model diagnostics 3, 1}
library(gridExtra)
grid.arrange(plot(ARMA.31, type=c("p","smooth")),
             plot(ARMA.31, sqrt(abs(resid(.)))~fitted(.),
                                 type=c("p","smooth"),ylab=expression(sqrt(abs(resid)))),
                                       ## "sqrt(abs(resid(x)))"),
         plot(ARMA.31, resid(.,type="pearson")~month_id,
                  type=c("p","smooth")),
         qqnorm(ARMA.31,abline=c(0,1)))
```
```{r autocorrelation diagnostic, 3,1}
grid.arrange(plot(ACF(ARMA.31),alpha=0.05),
             plot(ACF(ARMA.31,resType="normalized"),alpha=0.05),
             nrow=1)

acf(resid(ARMA.31))
pacf(resid(ARMA.31))
```

# in paper
```{r 'ARMA with p = 3, q = 0'}
ARMA.30 <- update(model.lme.ARMA, correlation = corARMA(p = 3, q = 0))
AIC(ARMA.30)

summary(ARMA.30)
Anova(ARMA.30)

plot(ARMA.30)

anova(ARMA.30, ARMA.31)
```

```{r autocorrelation model diagnostics 3, 0}
library(gridExtra)
grid.arrange(plot(ARMA.30, type=c("p","smooth")),
             plot(ARMA.30, sqrt(abs(resid(.)))~fitted(.),
                                 type=c("p","smooth"),ylab=expression(sqrt(abs(resid)))),
                                       ## "sqrt(abs(resid(x)))"),
         plot(ARMA.30, resid(.,type="pearson")~month_id,
                  type=c("p","smooth")),
         qqnorm(ARMA.30,abline=c(0,1)))
```

```{r autocorrelation diagnostic p3 q0}
grid.arrange(plot(ACF(ARMA.30),alpha=0.05),
             plot(ACF(ARMA.30,resType="normalized"),alpha=0.05),
             nrow=1)

acf(resid(ARMA.30))
pacf(resid(ARMA.30))

```

```{r ave temp by season and method}
library(emmeans)
pairs(emmeans(ARMA.30, ~ season * Method))

boxplot(aveTemp~season:Method, data = log_crw_month, las = 2)
boxplot(aveTemp~month_id, data = log_crw_month)

log_crw_month %>% 
   group_by(season, Method) %>% 
   dplyr::summarize(ave = mean(aveTemp), SD = sd(aveTemp), SE = SD/sqrt(n()))

library(lsmeans)
lsmeans(ARMA.30, pairwise ~ season * Method, adjust = 'tukey')

```

```{r season x method inteaction plot}
#log_crw_month$Method <- relevel(log_crw_month$Method, ref = 'CRWTL')

# reorder seasons
levels(log_crw_month$season)
log_crw_month$season <- factor(log_crw_month$season, levels = levels(log_crw_month$season)[c(3,1,4,2)])

library(rcompanion)
Sum_both = groupwiseMean(aveTemp ~ season * Method ,
                    data   = log_crw_month,
                    conf   = 0.95,
                    digits = 3,
                    traditional = FALSE,
                    percentile  = TRUE)
Sum_both

ggplot(Sum_both, aes(x =   season,
                y =    Mean,
                color = Method)) +
    geom_errorbar(aes(ymin=Percentile.lower,
                      ymax=Percentile.upper),
                   width=.2, size=0.7) +
    geom_point(shape=15, size=4) +
    theme_classic() +
    theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12)) +
    ylab("Mean temperature [C]") + xlab('Austral Season') +
    geom_hline(yintercept = 29.5, color = 'light blue', linetype = 2) +
    geom_hline(yintercept = 30.5, color = 'light blue') 

ggsave(here('figures/Fig7_anova_season-method_plot.jpg'), width = 7, height = 4)
```
