---
title: "Timor-Leste 2015-2017 temperature data"
author: "Catherine Kim"
date: "March 4th 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
```

Timor-Leste in situ and satellite temperature data from Nov 2015 to Jul 2017

Monthly average of daily maximum temperature

```{r combine  sites 5 and 10 m depth montly mean from daily max, crw data}
temp <- read_csv(here::here("data/alltemp_TL_data.csv")) %>% 
   mutate_if(is.character, as.factor)
head(temp)

temp$m.y <- as.factor(paste(temp$Month, temp$Year, sep = "/")) #rewrite m.y as month/year
temp$Month <- as.factor(temp$Month)
temp$Year <- as.factor(temp$Year)
temp$Depth <- as.factor(temp$Depth)
temp$Depth <- relevel(temp$Depth, ref = '5m')

# relevel Month/yr in correct order
levels(temp$m.y)

temp$m.y = factor(temp$m.y,levels(temp$m.y)[c(21,24,3,5,7,1,9,11,13,15,17,19,22,25,4,6,8,2,10,12,14,16,18,20,23)])
```

```{r boxplot Monthly temp ave by site}
ggplot(temp, aes(x = Site, y = Temp, fill = Year)) +
   facet_grid(Depth~.)+
    geom_boxplot() +
   geom_point(aes(color = Year), position = position_jitter(0.21)) +
    theme_classic()
```

Assigning the austral season to each Month 1-3 = summer etc.

```{r filter overlapping Crw and c data, add season, add method}
season <- data.frame(Month = 1:12, season = rep(c("summer", "fall", "winter", "spring"), each = 3))
season$Month <- as.factor(season$Month)
season

method <- data.frame(Site = c("Rural-N", "Urban-E", "Urban-W", "CRWTL"), Method = c("insitu", "insitu", "insitu", "sat"))
method

# subset out data with both CRW and logger temperature data
temp.same <- temp %>% filter(m.y != "8/2015" & m.y != "9/2015" & m.y != "10/2015" & m.y != "11/2015" & m.y != "6/2017" & m.y != "7/2017" & m.y != "8/2017") %>% 
  inner_join(season, by = 'Month') %>% 
   inner_join(method, by = 'Site') %>% 
   mutate_if(is.character, as.factor)

# remove CRW for one depth used for plot
temp.same <- temp.same[-which(temp.same$Site == 'CRWTL' & temp.same$Depth == '10m'),] #

# relevel method so sat is reference
temp.same$Method <- relevel(temp.same$Method, ref = "sat")

head(temp.same)
```

```{r crw}
crw <- temp %>% 
   filter( Site == "CRWTL") %>% 
   dplyr::select(Site, m.y, Temp, sd, se)
colnames(crw) <- c("Site", "m.y", "siteave", "SD", "SE")
crw
```

```{r average difference between 5 and 10 m temp loggers}
temp.same %>% 
   filter(Method != "sat") %>% 
   dplyr::select(-N, -sd, -se, -ci) %>% 
   pivot_wider(names_from = Depth,
               values_from = Temp) %>% 
   mutate(diff = `5m` - `10m`) %>% 
   group_by(Site) %>% 
   summarize(avediff = mean(diff)) %>% 
   summarize(diff = mean(avediff), SE = sd(avediff))
```

## Graph without facet by depth
average the 2 loggers per site
1 satellite point per Month

```{r ave temp loggers by site}
temp.avelog <- temp.same %>% 
   filter(Method != "sat") %>% 
   group_by(Site, m.y) %>% 
   dplyr::summarise(siteave = mean(Temp), SD = sd(Temp), SE = SD/sqrt(n())) #, N = n())

temp.avelog

alltemp <- rbind(temp.avelog, crw)
alltemp
```

```{r line graph all sites by Month/yr no depth facet}
start <-  c("11/2015", "11/2016")
end <-  c("5/2016", "5/2017")

cbPalette <- c("#D55E00", "#56B4E9", "#009E73",  "#CC79A7")

ggplot(alltemp, mapping= aes(m.y, siteave, group = Site, color = Site)) +
   geom_line()+
   geom_ribbon(aes(m.y, siteave, ymax= siteave+(SE), ymin= siteave-(SE), fill = Site),
               alpha = 0.4) + 
   scale_color_manual( values = cbPalette, name = "Site/Method") +
   scale_fill_manual(values = cbPalette, name = "Site/Method") +
   theme_classic() +
   xlab("Month/Year") + ylab("Temperature [C]") +
   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
         text = element_text(size = 14)) +
   geom_hline(yintercept = 29.5, color = "light blue", linetype = 2) +
   geom_hline(yintercept = 30.5, color = "light blue")  +
   annotate("rect", xmin = '11/2015', xmax = '5/2016', ymin = -Inf, ymax = Inf,
           alpha = .1,fill = "gray10") +
   annotate("rect", xmin = '11/2016', xmax = '5/2017', ymin = -Inf, ymax = Inf,
           alpha = .1,fill = "gray10")

ggsave(here::here("figures/Fig7_line_TLtemp.jpg"), width = 7, height = 4)
```

## Statistics

```{r histograph Temp, anova}
hist(temp.same$Temp)
```

testing site(3 logger sites and satellite) by season
```{r site by season anova}
site.lm <- lm(Temp~Site * season, data = temp.same)
library(car)
Anova(site.lm)
library(emmeans)
emmeans(site.lm, list(pairwise ~ season), adjust = "tukey")
emmeans(site.lm, list(pairwise ~ Site), adjust = "tukey")
```

```{r}
site.lm2 <- lm(Temp~Site * season, data = temp.same, contrasts = list(Method = contr.sum, season = contr.sum))
library(car)
Anova(site.lm2, type = 'III')
library(emmeans)
emmeans(site.lm2, list(pairwise ~ season), adjust = "tukey")
emmeans(site.lm2, list(pairwise ~ Site), adjust = "tukey")
```

```{r box plot showing method:season interaction}
ggplot(temp.same, aes(season, Temp)) + 
   geom_jitter(aes(color = Site)) +
   geom_boxplot(fill = NA) + 
   theme_classic()
```

testing method (logger or satellite) by season

```{r method by season anova}
method.lm <- lm(Temp~Method*season, data = temp.same)
Anova(method.lm)

emmeans(method.lm, list(pairwise ~ season), adjust = "tukey")
emmeans(method.lm, list(pairwise ~ Method), adjust = "tukey")
emmeans(method.lm, list(pairwise ~ Method:season), adjust = "tukey")

boxplot(Temp~Method, data = temp.same)
boxplot(Temp~season, data = temp.same)
boxplot(Temp~Method:season, data = temp.same, las = 2)
```

ANOVA with type III sums of sqs following https://www.r-bloggers.com/2011/03/anova-%E2%80%93-type-iiiiii-ss-explained/

```{r Type III SS method by season anova}
method.lm <- lm(Temp~Method*season, data = temp.same, contrasts = list(Method = contr.sum, season = contr.sum))
Anova(method.lm, type = 'III')

emmeans(method.lm, list(pairwise ~ season), adjust = "tukey")
emmeans(method.lm, list(pairwise ~ Method), adjust = "tukey")
emmeans(method.lm, list(pairwise ~ Method:season), adjust = "tukey")

boxplot(Temp~Method:season, data = temp.same, las = 2)

library(ggeffects)
plotdat <- method.lm %>% 
   ggemmeans(~ Method:season) %>% 
   as.data.frame() 

levels(plotdat$group)
plotdat$group <- ordered(plotdat$group, 
                         levels = c("summer", "fall", "winter", "spring"))

ggplot(plotdat, aes(x = group, y = predicted, color = x)) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                 width = 0.2, size = 1) +
   geom_point(shape = 15, size = 3) +
   theme_classic() +
   theme(text = element_text(size = 14)) +
   scale_x_discrete(breaks = c("summer", "fall", "winter", "spring"),
                    labels = c("Summer", "Fall", "Winter", "Spring")) +
   scale_color_discrete(name = "Method",
                        breaks = c("sat", "insitu"),
                        labels = c("CRWTL", "In Situ")) +
   xlab("Austral Season") +
   ylab("Temperature [C]") +
   geom_hline(yintercept = 29.5,
              linetype = "dashed",
              color = "lightblue") +
   geom_hline(yintercept = 30.5,
              color = "lightblue")

ggsave(here::here("figures/Fig8_seasonmethod.jpg"), 
       height = 4, width = 7, dpi = 300)
```

```{r box plot showing season }
ggplot(temp.same, aes(season, Temp)) + 
   geom_boxplot(fill = NA) + 
   geom_jitter(aes(color = Method)) +
   theme_classic()
```

```{r ave temp by CRW and all loggers together}
temp.same %>% 
   group_by(Method, season) %>% 
   dplyr::summarize(avetemp = mean(Temp), 
                    SDtemp = sd(Temp), 
                    SEtemp = sd(Temp)/sqrt(n()))
```

## Statistics testing site/depth differences between logger data

```{r}
logger <- filter(temp.same, temp.same$Method != 'sat') %>% droplevels()
head(logger)
```

```{r Anova on logger temperature differences between site, depth, Month/Year}
log.lm <- lm(Temp ~ Site * Depth + factor(Month)/factor(Year), data = logger)
Anova(log.lm)
plot(log.lm)

emmeans(log.lm, list(pairwise ~ factor(Month):factor(Year)), adjust = "tukey")
emmeans(log.lm, list(pairwise ~ Site:Depth), adjust = "tukey")

boxplot(Temp~Depth, data = logger)
boxplot(Temp~Site, data = logger)
boxplot(Temp~Month, data = logger)

logger$Ym <- interaction(logger$Year, logger$Month)
boxplot(Temp~Ym, data = logger, las = 2)
```

```{r Anova on logger temperature differences between site, Month/Year}
log.lm.nodepth <- lm(Temp ~ Site + factor(Month)/factor(Year), data = logger)
Anova(log.lm.nodepth)
plot(log.lm.nodepth)

emmeans(log.lm.nodepth, list(pairwise ~ factor(Month):factor(Year)), adjust = "tukey")
emmeans(log.lm.nodepth, list(pairwise ~ Site), adjust = "tukey")

boxplot(Temp~Site, data = logger)
boxplot(Temp~Month, data = logger)

logger$Ym <- interaction(logger$Year, logger$Month)
boxplot(Temp~Ym, data = logger, las = 2)
```

```{r }
sitedepth.ave <- logger %>% 
   group_by(Site,Depth) %>% 
   dplyr::summarize(ave = mean(Temp), SD = sd(Temp), se = SD/sqrt(n()))
sitedepth.ave

logger %>% 
   group_by(Month, Year) %>% 
   dplyr::summarize(ave = mean(Temp), SD = sd(Temp), se = SD/sqrt(n()))
```

```{r ave diff betwen site ave temp 5 vs 10 m}
sitedepth.diff <- sitedepth.ave %>% 
   select(Site, Depth, ave) %>% 
   spread(Depth, ave) %>% 
   mutate(diff = `10m` - `5m`)
sitedepth.diff
ave(sitedepth.diff$diff)
```