---
title: "Statistics on 2015-2017 coral health data from Timor-Leste"
author: "Catherine Kim"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## read in transect data with 8 categories including diseases white syndrome (WS) and growth anomalies (GAs)

```{r read in data with growth anomaolies (GAs) separated }
library(here)

compd.t <- read.csv(here::here('data/comp8GA_tave.csv'))
head(compd.t)
```

```{r create unique transect id}
compd.t$long_id <- as.factor(paste(compd.t$Site2, compd.t$Depth, compd.t$Transect, sep = "_"))

#write.csv(levels(compd.t$long_id), here("data/unique_ids.csv"))

# read file with num id
ids <- read.csv(here::here("data/unique_ids.csv"))

library(dplyr)
compd.t <- inner_join(compd.t, ids, by = 'long_id') %>% # join unique transect ids - num_id
   mutate_if(is.character, as.factor)
   # change num_id an year to factors

# change integers to factors as well
compd.t$num_id <- as.factor(compd.t$num_id)
compd.t$Year <- as.factor(compd.t$Year)

levels(compd.t$Site2)

glimpse(compd.t)
```

# Healthy coral statistics

```{r summarize healthy corals}
mean(compd.t$Healthy, na.rm = TRUE) # overall percent healthy coral
sd(compd.t$Healthy)/sqrt(length(compd.t)) # SE healthy coral

library(Rmisc)
comp.h <- summarySE(compd.t, "Healthy", c("Year", "Site2", "Depth"))
head(comp.h)

comp.hs <- summarySE(compd.t, "Healthy", c("Site2"))
head(comp.hs)
```

```{r healthy hist plot}
hist(compd.t$Healthy)
shapiro.test(compd.t$Healthy)

library(ggpubr)
ggboxplot(
  compd.t, x = "Site2", y = "Healthy", facet.by = 'Depth',
  color = "Year", palette = "jco"
  )
```

```{r healthy repeated measures}
library(nlme)
h.lme <- lme((Healthy) ~ Year * Site2 *Depth, random = ~1|num_id,
             data = compd.t,
             na.action = na.exclude)
h.lme
summary(h.lme)

library(car)
Anova(h.lme)  
library(emmeans)
emmeans(h.lme, list(pairwise ~ Year), adjust = "tukey")
emmeans(h.lme, list(pairwise ~ Site2), adjust = "tukey")
emmeans(h.lme, list(pairwise ~ Depth), adjust = "tukey")
```

## Disease/Compromised states statistics

```{r ave total compromised states}
# across site and depth, ave across years
compd.t %>% 
   mutate(comp = CCA + BL + AllAlgae + TL + OTH) %>% 
   group_by(Site2, Depth) %>% 
   dplyr::summarize(avecomp = mean(comp), SE = sd(comp)/sqrt(n()))

# overall compromised health average
compd.t %>% 
   mutate(comp = CCA + BL + AllAlgae + TL + OTH) %>% 
   dplyr::summarize(avecomp = mean(comp), SE = sd(comp)/sqrt(n()))
```

```{r total disease ave by site and depth}
library(tidyr)
library(dplyr)

# ave disease across years at site and depth
compd.t %>% 
   mutate(disease = WS + GA) %>%
   group_by(Site2, Depth ) %>% 
   dplyr::summarize(dave = mean(disease),
             SE = sd(disease)/sqrt(n()))

# proportion of disease at Rural-N
compd.t %>% 
   dplyr::mutate(disease = WS + GA)%>% 
   dplyr::summarize(sum(disease)) # total proportion of disease

compd.t %>% 
   dplyr::mutate(disease = WS + GA)%>% 
   filter(Site2 == 'Rural-N') %>% 
   dplyr::summarize(sum(disease)) # disease at Rural-N

12.04151	/ 26.84603	
```

```{r prevalence WS and GAs by year}
ws.ga.yr <- compd.t %>% 
   group_by( Year) %>% 
   dplyr::summarize(WSave = mean(WS), WSSE = sd(WS)/sqrt(n()),
             GAave = mean(GA), GASE = sd(GA)/sqrt(n()))
ws.ga.yr
```

```{r prevalence WS and GAs by site and year}
ws.ga.styr <- compd.t %>% 
   group_by(Site2, Year) %>%  
   dplyr::summarize(WSave = mean(WS), WSSE = sd(WS)/sqrt(n()),
             GAave = mean(GA), GASE = sd(GA)/sqrt(n()))
ws.ga.styr
```

```{r ave disease and comp health by site year}
cd <- compd.t %>% 
   dplyr::mutate(compd = CCA + BL + AllAlgae + TL + OTH + GA + WS)%>% 
   group_by(Site2,Year) %>% 
   dplyr::summarize(avecd = mean(compd),
             SE = sd(compd)/sqrt(n()))
cd
```

## ANOVAS of compromised health categories 

### bleaching

```{r bleaching hist and plot}
hist(compd.t$BL)
qqnorm(compd.t$BL)
#qqplot(compd.t$BL)

hist(sqrt(compd.t$BL))
qqnorm(sqrt(compd.t$BL))
#qqplot(sqrt(compd.t$BL))

library(car)
leveneTest((BL)~Year*Site2*Depth, compd.t)
```

```{r bleaching lmer repeated anova}
library(nlme)
bl.lme <- lme(sqrt(BL) ~ Year * Site2*Depth, random = ~1|num_id, data = compd.t,
              na.action = na.exclude)
bl.lme
summary(bl.lme)
library(car)
Anova(bl.lme) # year:site:depth interaction
library(emmeans)
bl.tuk <- emmeans(bl.lme, list(pairwise ~ Year:Site2:Depth), adjust = "tukey")
bl.tukpair <- as.data.frame(bl.tuk$`pairwise differences of Year, Site2, Depth`)
bl.tukpair[which(bl.tukpair$p.value < 0.05),]
```

```{r bleaching summary stats and plot}
summarySE(compd.t, "BL", c("Year", "Site2", "Depth"))

summarySE(compd.t, "BL", c("Year", "Site2"))

library(ggpubr)
ggboxplot(
  data = compd.t, x = "Site2", y = "BL", facet.by = 'Depth',
  color = "Year", palette = "jco"
  )
```

### algae overgrowth anova

```{r anova algae overgrowth}
hist(compd.t$AllAlgae + compd.t$CCA)
qqnorm(compd.t$AllAlgae + compd.t$CCA)
qqnorm(compd.t$AllAlgae)
leveneTest((AllAlgae)~Year*Site2*Depth, compd.t)

ggboxplot(
  data = compd.t, x = "Site", y = "AllAlgae", 
  color = "Year", palette = "jco"
  )
aa.se <- summarySE(compd.t, "AllAlgae", c("Year", "Site"))
aa.se
```

```{r algae lmer repeated anova}
aa.lme <- lme((AllAlgae) ~ Year * Site2 * Depth, random = ~1|num_id, data = compd.t, na.action = na.exclude)
aa.lme
summary(aa.lme)
Anova(aa.lme)  # Year:site interaction
emmeans(aa.lme, list(pairwise ~ Year:Site2), adjust = "tukey")
```

```{r ave all algae growth per site}
alg <- compd.t %>% 
   group_by(Site2, Year) %>% 
   dplyr::summarize(avealg = mean(AllAlgae),
             SE = sd(AllAlgae)/sqrt(n()))
alg
```

### CCA overgrowth

```{r CCCA}
hist(compd.t$CCA)
N <- length(compd.t)
compd.t$slogCCA <- log10((compd.t$CCA * (N-1) + .5)/N)
hist((compd.t$slogCCA))
#qqnorm(log10(compd.t$CCA+1))
qqnorm(sqrt(compd.t$CCA))

leveneTest(sqrt(CCA)~Year * Site2 * Depth, compd.t) # not significant
```
```{r cca summary}
cca.se <- summarySE(compd.t, "CCA", c("Year", "Site"))

ggboxplot(
  data = compd.t, x = "Site", y = "CCA", facet.by = 'Depth',
  color = "Year", palette = "jco"
  )
```

```{r CCA lmer repeated anova}
cca.lme <- lme((CCA) ~ Year * Site2 * Depth, random = ~1|num_id, data = compd.t,
             na.action = na.exclude)
cca.lme
summary(cca.lme)
Anova(cca.lme)  # sig year:site, site:depth interactions
emmeans(cca.lme, list(pairwise ~ Year:Site2), adjust = "tukey")
emmeans(cca.lme, list(pairwise ~ Site2:Depth), adjust = "tukey")
```

### Tissue loss (TL) ANOVA

```{r anova tissue loss}
hist(compd.t$TL)
#hist(log10(compd.t$TL+1))
hist(sqrt(compd.t$TL))
qqnorm(sqrt(compd.t$TL))
leveneTest((TL)~Year*Site2*Depth, compd.t) # not significant

ggboxplot(
  data = compd.t, x = "Site2", y = "TL", facet.by = 'Depth',
  color = "Year", palette = "jco"
  )
```

```{r TL lmer repeated anova}
tl.lme <- lme((TL) ~ Year * Site2 * Depth, random = ~1|num_id, 
              data = compd.t,
             na.action = na.exclude)
tl.lme
summary(tl.lme)
Anova(tl.lme)  # sig year
emmeans(tl.lme, list(pairwise ~ Year), adjust = "tukey")
```

# Plot of disease and compromised health states

## combine GAs with OTHER for plot
```{r}
compd.plot <- compd.t

compd.plot$OTH <-  compd.t$OTH + compd.t$GA
compd.plot$GA <- NULL
```

## average categories by site and convert to long format

```{r average disease and compromised health categories by site}
ave_compd <- compd.plot %>% 
   pivot_longer(AllAlgae:WS,
                names_to = "Category",
                values_to = "Percent") %>% 
   group_by(Year, Site2, Category) %>% 
      dplyr::summarize(avePer = mean(Percent))

ave_compd
```

## stacked graph of disease (only WS) and compromised states

```{r average by year and site}
ave_compd$Category <- factor(ave_compd$Category, levels=c("WS", "OTH", "TL", "CCA","AllAlgae", "BL", "Healthy"))

ave_compd
``` 

```{r no facet by depth}
ggplot(ave_compd, aes((Year), avePer,  fill = Category)) +
   facet_grid(.~Site2, switch = "x") +
    geom_bar(position="stack", stat="identity")+ 
 #  ggtitle("Signs of Compromised Health 2015, 2017") +
   theme_classic() +
   labs(x = "", y = "Percent") +
   theme(plot.title = element_text("sans", "bold",
                                   color = "black", 15)) +
   theme(text = element_text(size = 16)) +
   theme(axis.ticks.x = element_blank()) + #, panel.background = element_rect("gray85"))+
   theme(axis.text.x = element_text(hjust = 0.5),
         legend.title = element_blank())+
   scale_fill_brewer(palette="RdYlBu",
                     labels = c("WS", "OTH", "TL",
                                "CCA", "All Algae",
                                "BL", "Healthy"))

ggsave(here::here('figures/stacked-comphealth-20152017-nodepth.jpg'), width = 7, height = 4)
```