---
title: "Disease and comproised health data organization"
author: "CKim"
date: "August 7, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(Rmisc)
```


```{r read in TL15-health master}
library(readr)
library(here)
dat <- read_csv(here::here("data/CoralHealth-2015-2017edit042021-R-genmorph.csv"))

str(dat)
dat$Site <- as.factor(dat$Site)
levels(dat$Site)
dat$Number <- as.numeric(dat$Number)
dat$Depth <- as.factor(dat$Depth)
dat$Transect <- as.factor(dat$Transect)
dat$Status <- as.factor(dat$Status)

dat$GEN.morph <- as.factor(paste(dat$GEN, dat$morph, sep = ""))
levels(dat$GEN.morph)
head(dat)
```

```{r }
levels(dat$H1)
levels(dat$GEN)
levels(dat$Status)
dat[which(dat$H1==" "),]
colnames(dat)
```

```{r group by status, calculate prevalence}
library(tidyr)
library(dplyr)
head(dat)
dat.stat <- dat %>% group_by(Year, Site, Depth, Transect, Status) %>%
   dplyr::summarize(tot.stat = sum(Number)) %>%
   spread(Status, tot.stat, fill = 0) %>%
   mutate(tot.coral = Healthy + Comp + Disease + Bleaching + Pale)
head(dat.stat)

# include pale with bleaching
dat.pstat <- dat.stat %>%
   transmute(pBleaching = (Bleaching + Pale)/tot.coral * 100, pComp = Comp/tot.coral*100, pHealthy = Healthy/tot.coral*100,
             pDisease=Disease/tot.coral*100) %>%
   gather(Status, Percent, pBleaching, pComp, pHealthy, pDisease)

dat.pstat$Status <- as.factor(dat.pstat$Status)
head(dat.pstat)

w.site <- spread(dat.pstat, Status, Percent) %>% 
   mutate(total = pBleaching + pComp + pDisease + pHealthy)
(w.site)
```
```{r condition ave by site, depth}
stat.site <- summarySE(dat.pstat, measurevar = "Percent", groupvars= c("Year", "Site", "Depth", "Status"))
head(stat.site)
```

```{r graph status, echo = FALSE}
library(ggplot2)
stat.site$Depth <- relevel(stat.site$Depth, ref = "5m")

ggplot(stat.site, aes(factor(Year), Percent, fill = Status)) +
          facet_grid(Depth ~ Site) +
    geom_bar(position="stack",stat="identity") +
       ggtitle("Coral Status 2015 and 2017") +
   theme_classic()
```

## total disease counts
```{r total disease summed by site depth year}
# TRE
dat[which(dat$H2 == 'TRE'),]

distot <- dat %>% 
   group_by(Year, Site,  Status) %>% 
   dplyr::summarize(total = sum(Number))
distot

distotgen <- dat %>% 
   group_by(Year, Site, Status, GEN, morph,H1) %>% 
   dplyr::summarize(total = sum(Number)) %>% 
   filter(Status == 'Disease')
distotgen

# porportion of WS in Beloi 2015
13/21
```
```{r WS summed by site depth year}
WStotgen <- dat %>% 
   group_by(Year, GEN, Site, morph, Status, H1, H2) %>% 
   dplyr::summarize(total = sum(Number))%>% 
   filter(H1 == "WS" | H2 == 'WS')
WStotgen
sum(WStotgen$total)
```
```{r GA summed by gen, site, depth, year}
GAtotgen <- dat %>% 
   group_by(Year, Site, GEN, morph,H1, H2, Status) %>% 
   dplyr::summarize(total = sum(Number)) %>% 
  filter(H1 == "GA" | H2 == 'GA') %>% 
   droplevels()
GAtotgen
sum(GAtotgen$total)
levels(GAtotgen$GEN)  # 9 genera with GAs

TLtotgen <- dat %>% 
   group_by(Year, Site, GEN, morph,H1, H2, Status) %>% 
   dplyr::summarize(total = sum(Number)) %>% 
  filter(H1 == "TL" | H2 == 'TL') %>% 
   droplevels()
TLtotgen
sum(TLtotgen$total)

BLtotgen <- dat %>% 
   group_by(Year, Site, GEN, morph,H1, H2, H3, H4, Status) %>% 
   dplyr::summarize(total = sum(Number)) %>% 
  filter(Status == 'Bleaching') %>% 
   droplevels()
BLtotgen
sum(BLtotgen$total)
```
```{r subset bleaching}
dat.blp <- dat %>% filter( dat$Status == "Bleaching")
colnames(dat.blp)[which(colnames(dat.blp) == "H1")] <- "H"
nrow(dat.blp)
head(dat.blp)
 dat.blp$H2 <- NULL
 dat.blp$H3 <- NULL
 dat.blp$H4 <- NULL

# not necessary to include H2, all included when subsetting bleaching, will throw off percent totals
# dat.bl2 <- filter(dat.blp, dat.blp$H2 != "")
# 
# colnames(dat.bl2)[which(colnames(dat.bl2) == "H2")] <- "H"
# head(dat.bl2)
# nrow(dat.bl2)
# 
# dat.blall <- rbind(dat.blp, dat.bl2)

# head(dat.blall)
# nrow(dat.blall)
```
```{r subset diseae}
dat.d <- dat %>% filter(dat$Status == "Disease")
nrow(dat.d)
head(dat.d)
colnames(dat.d)[which(colnames(dat.d) == "H1")] <- "H"

 dat.d$H2 <- NULL
 dat.d$H3 <- NULL
 dat.d$H4 <- NULL

# not necessary to inclue H2 column all subset with status = disease
# dat.d2 <- filter(dat.d, dat.d$H2 != "")
# dat.d2
# nrow(dat.d2)
# 
# dat.dall <- rbind(dat.d, dat.d2)

# nrow(dat.dall)
# head(dat.dall)
```
```{r subset compromised }
dat.comp <- dat %>% filter(dat$Status == "Comp")
head(dat.comp)
nrow(dat.comp)

dat.comph2 <- filter(dat.comp, dat.comp$H2 != "")

dat.comph3 <- filter(dat.comp, dat.comp$H3 != "")

dat.comph4 <- filter(dat.comp, dat.comp$H4 != "")
dat.comp[which(dat.comp$H4 == "Broken"),]

dat.comp$H2 <- NULL
dat.comp$H3 <- NULL
dat.comp$H4 <- NULL
colnames(dat.comp)[which(colnames(dat.comp) == "H1")] <- "H"

dat.comph2$H1 <- NULL
dat.comph2$H3 <- NULL
dat.comph2$H4 <- NULL
colnames(dat.comph2)[which(colnames(dat.comph2) == "H2")] <- "H"
head(dat.comph2)
nrow(dat.comph2)

dat.comph3$H1 <- NULL
dat.comph3$H2 <- NULL
dat.comph3$H4 <- NULL
colnames(dat.comph3)[which(colnames(dat.comph3) == "H3")] <- "H"
head(dat.comph3)
nrow(dat.comph3)

dat.comph4$H1 <- NULL
dat.comph4$H2 <- NULL
dat.comph4$H3 <- NULL
colnames(dat.comph4)[which(colnames(dat.comph4) == "h4")] <- "H"
head(dat.comph4)
nrow(dat.comph4)

dat.compall <- rbind(dat.comp, dat.comph2, dat.comph3, dat.comph4, dat.d, dat.blp)
head(dat.compall)
nrow(dat.compall) 

nrow(dat.comph3) + nrow(dat.comph2) +nrow(dat.comph4) + nrow(dat.comp)
nrow(dat.d)
nrow(dat.blp)

levels(dat.compall$H)
dat.compall[which(dat.compall$H == "cyl"),]

```

```{r total corals per transect}
dat.compall.sum <- dat.compall %>% group_by(Year, Site, Depth, Transect, H) %>%
   dplyr::summarize(tot.stat = sum(Number)) %>%
   spread(H, tot.stat, fill = 0)
head(dat.compall.sum)
str(dat.compall.sum)

# add compall total to sum of compromised stats - no repeats - to compare
compall.tot <- apply(dat.compall.sum[,5:ncol(dat.compall.sum)], 1, sum)

x <- data.frame(1:48)
head(x)

for( i in 5:ncol(dat.compall.sum)){
   y <- dat.compall.sum[,i]/compall.tot*(w.site$pBleaching+w.site$pComp+w.site$pDisease) # scale to percent withouth h2, h3, h4 cols
   x <- cbind(x,y)
}
head(x)
str(x)

dat.compall.per <- cbind.data.frame(dat.compall.sum[,1:4], x, w.site$pHealthy)
dat.compall.per$X1.48 <- NULL
dat.compall.per$`w.site$pHealthy`
colnames(dat.compall.per)[which(colnames(dat.compall.per) == "w.site$pHealthy")] <- 'Healthy'

totpercent <- dat.compall.per %>% dplyr::summarize(total = sum(5:ncol(dat.compall.per)))

head(dat.compall.per)

compall.per.nobd10 <- dat.compall.per %>% filter(Site != "Beloi") %>%
   filter(Site != "DiliRock" | Depth == "5m")
#write.csv(dat.compall.per, "TL15-allcompcats-per-wide.csv")

# above code works better
## calc percent using total transect
# trans.tot <- dat %>% 
#    group_by(Year, Site, Depth, Transect) %>% 
#    dplyr::summarise(tot = sum(Number))
# 
# z <- data.frame(1:48)
# 
# for( i in 5:ncol(dat.compall.sum)){
#    y <- dat.compall.sum[,i]/trans.tot$tot*100
#    z <- cbind(z,y)
# }
# head(z)
# 
# dat.compall.per <- cbind.data.frame(dat.compall.sum[,1:4], z)
# dat.compall.per$X1.48 <- NULL
```
```{r cyanoovergrowth by site and depth}
cyno <- dat.compall.per %>% 
   group_by(Year, Site, Depth) %>% 
   dplyr::summarize(aveCY = mean(CY), 
             SE = sd(CY)/sqrt(n()))
```

```{r filter out beloi - Dilirock 10m for no scaling with pcomp PBL pD}
# head(dat.compall)
# dat.compall.sum <- dat.compall %>% dplyr::group_by(Site, Depth, Transect, H) %>%
#    dplyr::summarize(tot.stat = sum(Number)) %>%
#    spread(H, tot.stat, fill = 0)
# head(dat.compall.sum)
# 
# bd10.compall <- dat.compall.sum %>% filter(Site == "Beloi" | Site == "DiliRock") %>%
#    filter(Site == "Beloi" | Depth == "10m")
# 
# compall.tot <- apply(dat.compall.sum[,4:ncol(dat.compall.sum)], 1, sum)
# 
# x <- data.frame(1:9)
# head(x)
# 
# for( i in 4:ncol(bd10.compall)){
#    y <- bd10.compall[,i]/compall.tot*100
#    x <- cbind(x,y)
# }
# head(x)
# str(x)
# 
# bd10.compall.per <- cbind.data.frame(bd10.compall[,1:3], x)
# bd10.compall.per$X1.9 <- NULL
# 
# head(bd10.compall.per)
# ```
# ```{r}
# compall <- rbind(bd10.compall.per, compall.per.nobd10)
# head(compall)
```

# 8 categories inc GAs
```{r read and join combined comproimsed categories - 8 categories inc GAs}
## checked with GA as separate category and averages less than 1% - grouped with other in compcat7
categories <- read.csv(here::here("data/status_levels_join.csv"))

comp8GA <- dat.compall.per %>% 
   pivot_longer(cols = 5:ncol(dat.compall.per), names_to = 'H', values_to = 'Percent')%>% 
   inner_join(categories, by = 'H') %>% 
   mutate_if(is.character, as.factor)%>% 
   group_by(Year, Site, Depth, Transect, Compcats8GA)%>% 
   dplyr::summarize(totcat = sum(Percent)) %>% 
   ungroup() 
head(comp8GA)

# wide format
comp8GA_t_wide <- pivot_wider(comp8GA, names_from = 'Compcats8GA', 
                            values_from = 'totcat')
head(comp8GA_t_wide)
# write file
write.csv(comp8GA_t_wide, here::here('data/comp8GA_tave.csv'), row.names = T)
```

```{r comp 8 cataegories w GA averages by site, depth, year}
comp8_yr_sited_ave <- comp8GA %>% 
   group_by(Year, Site, Depth, Compcats8GA)%>% 
   dplyr::summarize(aveper = mean(totcat))

head(comp8_yr_sited_ave)

# check total sum of transect = 100
 check_comp8 <- comp8_yr_sited_ave %>% 
    pivot_wider(names_from = 'Compcats8GA', values_from = 'aveper')%>% 
    mutate(tot = BL + OTH + TL + CCA + AllAlgae + WS +  Healthy) 
 View(check_comp8)
```
# GAs <1% prevlence for all sites
# compromised categories inc healthy
```{r read and join combined comproimsed categories}
## checked with GA as separate category and averages less than 1% - grouped with other in compcat7


comp7.combo <- dat.compall.per %>% 
   pivot_longer(cols = 5:ncol(dat.compall.per), names_to = 'H', values_to = 'Percent')%>% 
   inner_join(categories, by = 'H') %>% 
   mutate_if(is.character, as.factor)%>% 
   group_by(Year, Site, Depth, Transect, Compcats7)%>% 
   dplyr::summarize(totcat = sum(Percent)) %>% 
   ungroup() 
head(comp7.combo)

# wide format
comp7_t_wide <- pivot_wider(comp7.combo, names_from = 'Compcats7', 
                            values_from = 'totcat')
head(comp7_t_wide)
# write file
write.csv(comp7_t_wide, here::here('data/comp7_tave.csv'))
```
```{r comp 7 averages by site, depth, year}
comp7_yr_sited_ave <- comp7.combo %>% 
   group_by(Year, Site, Depth, Compcats7)%>% 
   dplyr::summarize(aveper = mean(totcat))

head(comp7_yr_sited_ave)

# check total sum of transect = 100
 check_comp7 <- comp7_yr_sited_ave %>% 
    pivot_wider(names_from = 'Compcats7', values_from = 'aveper')%>% 
    mutate(tot = BL + OTH + TL + CCA + AllAlgae + WS +  Healthy) 
 View(check_comp7)
```

```{r plot with first submission compcat7}
comp7_yr_sited_ave$Depth <- relevel(comp7_yr_sited_ave$Depth, ref = '5m') # 5m as reference for depth

# reorder compcat catetories Healthy on Bottom, WS on top

ggplot(comp7_yr_sited_ave, aes(factor(Year), aveper, fill = Compcats7)) +
          facet_grid(Depth ~ Site, switch = 'x') +
    geom_bar(position="stack", stat="identity") +
   ggtitle("Signs of Compromised Health 2015, 2017") +
   theme_classic()
```

## 7 compromised health categories - manually
```{r manual join of compromised health categories}
comp8.combo <- dat.compall.per[,1:4]

comp8.combo$BL <- dat.compall.per$BL + dat.compall.per$PBL + dat.compall.per$Pale +
   dat.compall.per$FBl + dat.compall.per$PBT + dat.compall.per$PB + dat.compall.per$Tip + dat.compall.per$UB #focal bleaching
 #no PBT

comp8.combo$OTH <- dat.compall.per$BAR + dat.compall.per$BARN +dat.compall.per$Broken + dat.compall.per$BUR + dat.compall.per$contact+
   dat.compall.per$Coral + dat.compall.per$Crabgall + dat.compall.per$DRU + dat.compall.per$FW + + dat.compall.per$INV +
   dat.compall.per$Mucus + dat.compall.per$PIG +
   dat.compall.per$PRED + dat.compall.per$`Pred?` + dat.compall.per$Pred + dat.compall.per$SED + 
   dat.compall.per$MILLE + dat.compall.per$Mucus +
   dat.compall.per$Soft + dat.compall.per$SoftCoral + dat.compall.per$SP + dat.compall.per$TUN + dat.compall.per$TUN + 
 dat.compall.per$TRE +
   dat.compall.per$upside + dat.compall.per$Xmas

comp8.combo$TL <- dat.compall.per$whspots+ dat.compall.per$TL + dat.compall.per$whspot # dat.compall.per$WH no longer in columns
   
comp8.combo$CCA <- dat.compall.per$CCA

comp8.combo$AllAlgae <- dat.compall.per$Chlorodesmis + dat.compall.per$CY +
   dat.compall.per$HAL + dat.compall.per$Hali + dat.compall.per$MA + dat.compall.per$turf +
   dat.compall.per$TUR

comp8.combo$WS <- dat.compall.per$WS

comp8.combo$GA <- dat.compall.per$GA

head(comp8.combo)

comp8.combo$Healthy <- w.site$pHealthy


comph8.combo <- mutate(comp8.combo, tot = BL + OTH + TL + CCA + AllAlgae + WS + GA +  Healthy)
head(comph8.combo)

comp8.lcombo <- gather(comp8.combo, Comp, Percent, BL, AllAlgae, OTH, CCA, TL, WS)
head(comp8.lcombo)
comp8.lcombo$Comp <- as.factor(comp8.lcombo$Comp)
levels(comp8.lcombo$Comp)
```
```{r}

## diseases, WS added to comp.lcombo
dis <- dat.compall.per[,1:3]
dis$TRE <- dat.compall.per$TRE
dis$WS <- dat.compall.per$WS
dis$GA <- dat.compall.per$GA

head(dis)

compdis <- cbind(comp.combo, dis[4:6])
head(compdis)
write.csv(compdis, "Percent-CompDisease-TL2015-sdtrans.csv")
```
```{r 7 compcats inc health percent}
comp15.7site <- summarySE(comp7.lcombo,  measurevar = "Percent", groupvars= c("Site", "Depth", "Comp"))
head(comp15.7site)
comp15.7site$Depth <- relevel(comp15.7site$Depth, ref = "5m")
colnames(comp15.7site)[which(colnames(comp15.7site) == "Comp")] <- "Status"

write.csv(comp15.7site, "TL15-compcats7-WS-site-summarySE.csv")

h.site <- filter(stat.site, Status == "pHealthy")
h.site$Status <- "Healthy"
head(h.site)

comph15.7 <- rbind(comp15.7site, h.site)
head(comph15.7)
write.csv(comph15.7, "TL15-comph7-WS-site-summarySE.csv")
```

```{r graph comp coniditions, echo = FALSE}
ggplot(comph15.7, aes(Site, Percent, fill = Status)) +
          facet_grid(Depth ~.) +
    geom_bar(position="stack", stat="identity") +
   ggtitle("Signs of Compromised Health 2015") +
   theme_classic()
```

## status by genera count ----
```{r}
dat.gen <- dat %>% group_by(Site, Depth, Status, GEN.morph, GEN) %>%
   summarize(tot.stat = sum(Number)) %>%
   spread(Status, tot.stat, fill = 0) %>%
   mutate(tot.coral = Healthy + Comp + Disease + Bleaching + Pale) %>%
   filter(Bleaching != 0 | Pale != 0)

dat.gen$Comp <- NULL
dat.gen$Healthy <- NULL
dat.gen$Disease <- NULL
head(dat.gen)

long.gen <- gather(dat.gen, Bleach, Number, Bleaching, Pale, factor_key = T)
long.gen
```
```{r seriatopora}
head(dat)
dat.seri <- dat %>% group_by(Site, Depth, Transect) %>%
   filter(GEN.morph == "SER") %>%
   group_by(Site, Depth, Status, Transect, GEN) %>%
   summarize(tot.stat = sum(Number)) %>%
   spread(Status, tot.stat, fill = 0) %>%
   mutate(tot.coral = Healthy + Comp  + Bleaching )
head(dat.seri)
```


```{r graph # bleaching genera by site}
long.gen$Depth <- relevel(long.gen$Depth, ref = "5m")
ggplot(long.gen,  aes(GEN.morph, Number, fill = Bleach)) +
   facet_grid(Depth ~Site) +
    geom_bar(position=position_dodge(), stat="identity") +
     theme(axis.text.x  = element_text(angle=90))

setwd("C:/Users/jk845/OneDrive/Documents/Timor-Leste/2015-2017-CoralData")
write.csv(long.gen, "TL2015-Bl-genera-sum.csv")

```

```{r subset by gen >1}
long.gen1 <- filter(long.gen, Number > 1)
long.gen1$GEN.morph <- as.factor(long.gen1$GEN.morph)
levels(long.gen1$GEN.morph)

ggplot(long.gen1,  aes(GEN.morph, Number, fill = Bleach)) +
   facet_grid(Depth ~Site) +
     geom_bar(position=position_dodge(), stat="identity") +
  theme(axis.text.x  = element_text(angle=90)) +
   ggtitle("Count of Colonies Bleaching by Site 2015")
```
   
## status by genera sites combined ----
```{r}
dat.gen.ns <- dat %>% group_by( Depth, Status, GEN.morph ) %>%
   summarize(tot.stat = sum(Number)) %>%
   spread(Status, tot.stat, fill = 0) %>%
   mutate(tot.coral = Healthy + Comp + Disease + Bleaching + Pale) %>%
   filter(Bleaching != 0 | Pale != 0)

dat.gen.ns$Comp <- NULL
dat.gen.ns$Healthy <- NULL
dat.gen.ns$Disease <- NULL
head(dat.gen.ns)

long.gen.ns <- gather(dat.gen, Bleach, Number, Bleaching, factor_key = T)

```

```{r graph # bleaching genera no site}

long.gen.ns$Depth <- relevel(long.gen$Depth, ref = "5m")
ggplot(long.gen.ns,  aes(GEN.morph, Number)) +
   facet_grid(Depth ~.) +
    geom_bar(position=position_dodge(), stat="identity") +
     theme(axis.text.x  = element_text(angle=90)) +
   ggtitle("Count of Bleaching Genera 2015")
```



